---
title: "HW 2"
author: "AJ Catalano"
date: "11/4/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Question 1

#### Reading and cleaning Mr. Trash Wheel data:

```{r}
trash_data =
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             range = "Mr. Trash Wheel!A2:N535") %>%
  filter(Dumpster != "NA") %>%
  mutate_at("Sports Balls", round) %>%
  janitor::clean_names() %>%
  rename(dumpster_id = dumpster)
```

#### Reading and cleaning 2018 and 2019 precipitation data:

```{r}
# 2018 data

precipitation_2018 =
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             range = "2018 Precipitation!A2:B14") %>%
  mutate(year = 2018) %>%
  relocate(year) %>%
  janitor::clean_names()

# 2019 data

precipitation_2019 =
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             range = "2019 Precipitation!A2:B14") %>%
  mutate(year = 2019) %>%
  relocate(year) %>%
  janitor::clean_names()
```

#### Combining 2018 and 2019 precipitation data:

```{r}
precipitation_data = 
  bind_rows(precipitation_2019, precipitation_2018) %>%
  mutate(month = month.name[month])
```

The Mr. Clean Wheel dataset contains information about the contents of `r nrow(trash_data)` dumpsters, including the weight (tons) and volume ($$yd^3$$) of trash and its type (e.g. plastic bottles, cigarette butts, and sports balls). **In 2019, the median number of sports balls collected in a dumpster was  `r trash_data %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.**

The Precipitation dataset contains `r nrow(precipitation_data)` monthly observations from 2018 - 2019 for total precipitation in inches. There are `r ncol(precipitation_data)` variables in the dataset. **The total precipitation in 2018 was `r precipitation_data %>% filter(year == 2018) %>% pull(total) %>% sum()`**

## Question 2

Reading and cleaning pols_month data:

```{r}
pols_data =
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(month = month.name[month],
         president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) %>%
         select(-prez_dem, -prez_gop, -day) %>%
  relocate(year, month, president)
```

Reading and cleaning snp data:

```{r}
snp_data =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  mutate(date = lubridate::mdy(date)) %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(year = ifelse(year > 2021, year - 100, year)) %>%
  arrange(year, month) %>%
  mutate(month = month.name[month]) %>%
  select(-day) %>%
  relocate (year, month)
```

Reading and cleaning unemployment data:

```{r}
unemployment_data = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>%
  janitor::clean_names() %>%
  mutate(
    month = factor(month, levels = month.abb)) %>%
  arrange(month, year) %>%
  mutate(month = recode(month,
             Jan = "January", Feb = "February",
             Mar = "March", Apr = "April",
             May = "May", Jun = "June",
             Jul = "July", Aug = "August",
             Sep = "September", Oct = "October",
             Nov = "November", Dec = "December"))
```

Merging SNP into Pols and then Unemployment:

```{r}
data_538 = 
  left_join(pols_data, snp_data) %>%
  left_join(., unemployment_data)
```

The SNP dataset includes monthly S&P 500 closing prices from `r snp_data %>% pull(year) %>% range()`. The Unemployment dataset contains the monthly unemployment rate from `r unemployment_data %>% pull(year) %>% range()`. The Pols dataset contains monthly entries on the political party affiliation of major federal and state elected officials from `r pols_data %>% pull(year) %>% range()`. The complete dataset, data_538, contains `r ncol(data_538)` variables and `r nrow(data_538)` observations.

## Question 3

```{r}
# if childs first name and ethnicity are not converted to sentence, the resulting table in the next code chunk fails to create columns for 2011 and 2012. Why?!

pop_baby_names = 
  read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    ethnicity = recode(ethnicity,
                       "ASIAN AND PACIFIC ISLANDER" = "Asian and Pacific Islander",
                       "BLACK NON HISPANIC" = "Black non Hispanic",
                       "HISPANIC" = "Hispanic",
                       "WHITE NON HISPANIC" = "White non Hispanic",
                       "ASIAN AND PACI" = "Asian and Pacific Islander",
                       "BLACK NON HISP" = "Black non Hispanic",
                       "WHITE NON HISP" = "White non Hispanic"),
    gender = recode(gender,
                    "FEMALE" = "Female",
                    "MALE" = "Male"),
    childs_first_name = str_to_sentence(childs_first_name),
    ethnicity = str_to_sentence(ethnicity)) %>%
  distinct() %>%
  arrange(year_of_birth, ethnicity, rank)
```

Making tables for the masses:

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year.

```{r}
pop_baby_names %>%
  filter(childs_first_name == "Olivia", gender == "Female") %>%
  select(ethnicity, year_of_birth, rank) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  )
```

Produce a similar table showing the most popular name among male children over time.

```{r}
pop_baby_names %>%
  filter(gender == "Male", rank == 1) %>%
  select(ethnicity, year_of_birth, childs_first_name) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name
  )
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
pop_baby_names %>%
  filter(gender == "Male",
         ethnicity == "White non hispanic",
         year_of_birth == 2016) %>%
  ggplot(
    aes(x = rank, y = count)) +
  geom_point() +
  theme_minimal()
```

